---
title: "DataAnalytics_assignment5"
format: html
editor: visual
---

# Assignment 5 - Generate Sampling Distributions and Confidence Intervals

## Challenge 1

### Step 1

```{r}
#Loading necessary libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(mosaic)
library(cowplot)

#Loading in the dataset as a tibble named 'd'
d <- read.csv("https://raw.githubusercontent.com/difiore/ada-datasets/main/IMDB-movies.csv")

```

### Step 2

```{r}
#Filtering the list and adding a new column
d2 <- d %>%
  filter(startYear >= 1920, startYear <= 1979, 
         runtimeMinutes >= 60 & runtimeMinutes <= 180) %>%
  mutate(decade = paste0((startYear %/% 10) * 10, "s"))

```

### Step 3

```{r}
#Using ggplot2 to plot histograms of the distribution of runtimes for each decade
ggplot(d2, aes(x = runtimeMinutes)) +
  geom_histogram(binwidth = 10, color = "pink", fill = "cyan") + #pretty colors
  facet_wrap(~ decade) #grouping by decade


```

### Step 4

```{r}
#Creating a new dataframe 'results' that contains the population mean and sd for each decade
results <- d2 %>% 
  group_by(decade) %>% 
  summarise(pop_mean = mean(runtimeMinutes, na.rm = TRUE), 
            pop_sd = sd(runtimeMinutes, na.rm = TRUE))
```

### Step 5 & Step 6 Combined (note se_mean)

```{r}
#Drawing a sample of 100 movies and calculating sample mean and sd of runtime
sample_runtime <- d2 %>%
   slice_sample(n = 100, replace = FALSE) %>% #pulling 100 movies without replacing
  summarise(
    s_mean = mean(runtimeMinutes), #mean of sample
    s_sd   = sd(runtimeMinutes), #sd of sample
    s_number   = n(), #confirmation of total number of samples
    se_mean = s_sd / sqrt(s_number), #standard error of the sample mean
    decade #adding decade column for inner_join in step 7 (whoops)
  )
```

### Step 7

```{r}
#Creating a new dataframe called comparison that joins population 'results' and sample_runtime 
comparison <- results %>%
  inner_join(sample_runtime, by = "decade") %>%
  mutate(
    diff_mean   = s_mean - pop_mean) %>%
  arrange(decade)
```

### Step 8

```{r}
#Running a sample distribution of mean runtime by each decade
#PReliminary
reps <- 1000  # number of replications
samp_dist <- vector("list", reps) #creating my dataframe as a vector list with my reps
#Bulk of code
for (reps in seq_len(i)) {
  samp_dist[[i]] <- d2 %>%
  #sampled_reps <- d2 %>% #designating reps to be pulled from d2
    group_by(decade) %>%
    group_modify(~ slice_sample(.x, n = 100, replace = FALSE)) %>% #using splice_sample function again to pull my samples
    summarise(
      s_mean = mean(runtimeMinutes, na.rm = TRUE), #sample mean
      s_sd   = sd(runtimeMinutes,   na.rm = TRUE) #sample sd
    )
}
samp_dist <- bind_rows(samp_dist) #switching from list to proper dataframe
```

### Step 9

```{r}
#Calculating the mean and sd of the sampling distribution of sample means (what a mouthful)
samp_summary <- samp_dist %>%
  group_by(decade) %>%
  summarise(
    s_dist_mean = mean(s_mean, na.rm = TRUE),
    samp_dist_sd = sd(s_mean, na.rm = TRUE)
  ) %>%
  arrange(decade)

#Creating a histogram for the samp_summary
ggplot(samp_dist, aes(x = s_mean)) + #I really am unsure if this is what is being asked but hopefully going to office hours might help
  geom_histogram(binwidth = 1, color = "#FF10F0", fill = "#00FFFF") +
  facet_wrap(~ decade)

```

### Step 10

```{r}
#Now comparing population runtime with first and second sample runtime
comparison2 <- results %>%
  inner_join(sample_runtime, by = "decade") %>%
  inner_join(samp_dist, by = "decade") %>%
  arrange(decade)
```

## Challenge 2

### Step 1

```{r}
#Loading in the zombie dataset as z
z <- read.csv("https://raw.githubusercontent.com/difiore/ada-datasets/main/zombies.csv")
```

### Step 2

```{r}
#Calculating the population mean and standard deviation 
zombie_results <- z %>% 
  summarise(height_mean = mean(height, na.rm = TRUE), 
            height_sd = sqrt(mean((height - mean(height))^2)),
            weight_mean = mean(weight, na.rm = TRUE),
            weight_sd = sqrt(mean((weight - mean(weight))^2)),
            age_mean = mean(age, na.rm = TRUE),
            age_sd = sqrt(mean((age - mean(age))^2)),
            num_kill_mean = mean(zombies_killed, na.rm = TRUE),
            num_kil_sd = sqrt(mean((zombies_killed - mean(zombies_killed))^2)),
            edu_mean = mean(years_of_education, na.rm = TRUE),
            edu_sd = sqrt(mean((years_of_education - mean(years_of_education))^2))
            )
```

### Step 3

\`\`\`{r} ggplot(z, aes(x = height)) + geom_boxplot() + facet_wrap(\~ gender)

```{r}
#Using ggplot2 to plot histograms of each of the variables above by gender

#Boxplot for height
height_box <- ggplot(z, aes(x = height)) +
  geom_histogram(binwidth = 10, color = "#FFF01F", fill = "#39ff14") + #not so pretty colors
  facet_wrap(~ gender) #grouping by gender

#Boxplot for weight
weight_box <- ggplot(z, aes(x = weight)) +
  geom_histogram(binwidth = 10, color = "#FFF01F", fill = "#39ff14") + #not so pretty colors
  facet_wrap(~ gender) #grouping by gender

#Boxplot for age
age_box <- ggplot(z, aes(x = age)) +
  geom_histogram(binwidth = 10, color = "#FFF01F", fill = "#39ff14") + #not so pretty colors
  facet_wrap(~ gender) #grouping by gender

#Boxplot for # Killed
kill_box <- ggplot(z, aes(x = zombies_killed)) +
  geom_histogram(binwidth = 10, color = "#FFF01F", fill = "#39ff14") + #not so pretty colors
  facet_wrap(~ gender) #grouping by gender

#Boxplot for education
edu_box <- ggplot(z, aes(x = years_of_education)) +
  geom_histogram(binwidth = 10, color = "#FFF01F", fill = "#39ff14") + #not so pretty colors
  facet_wrap(~ gender) #grouping by gender

#Placing all boxplots together
plot_grid(age_box, edu_box, height_box, kill_box, weight_box, ncol = 2)
```

### Step 4

```{r}
#Setting up gender as a factor for split colors
z$sex <- factor(z$gender)

#Scatterplot for height
ggplot(data = z, aes(x = age, y = height, color = gender)) + 
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  labs(x = "Age", y = "Height")


#Scatterplot for height
ggplot(data = z, aes(x = age, y = weight, color = gender)) + 
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  labs(x = "Age", y = "Weight")
```

There appears to be a positive correlation between age and height as well as age and weight, although the more apparent trends lies in the difference between gender.

### Step 5

```{r}
#Using Q-Q plots to assess distribution

#Variable 'age' distribution
qqnorm(z$age, main = "QQ Plot - Age Distribution", col = "#FFE5B4")
  qqline(z$age, col = "#FF0000")
#Normal distribution

#Variable 'weight' distribution
qqnorm(z$weight, main = "QQ Plot - Weight Distribution", col = "#FFE5B4")
  qqline(z$weight, col = "black")
#Suggests Normal Distribution

#Variable 'height' distribution
qqnorm(z$height, main = "QQ Plot - Height Distribution", col = "#FF0000")
  qqline(z$height, col = "black")
#End curvurture suggests difference in tail weight; validate with boxplot
ggplot(z, aes(x = height))
  geom_histogram(binwidth = 10, color = "black", fill = "red")
#Potential right-skew
library(moments)
skewness(z$height)
#Slightly right-skewed

#Variable 'zombies_killed' distribution
ggplot(z, aes(x = zombies_killed)) +
  geom_histogram(binwidth = 10, color = "red", fill = "pink")
skewness(z$zombies_killed)
#strong right skew

#Variable 'years_of_education' distribution
ggplot(z, aes(x = years_of_education)) +
  geom_histogram(binwidth = 10, color = "red", fill = "pink")
skewness(z$years_of_education)
#right skew

detach(package:moments)
```

### Step 6

```{r}
#height, weight, age, zombies_killed, years_of education


sample_z <- z %>%
   slice_sample(n = 50, replace = FALSE) %>% #pulling 50 survivors without replacing
  summarise(
    s_number = n(), #confirmation of total number of samples
    z_mean_height = mean(height), #mean of sample height
    z_sd_height = sd(height), #sd of sample height
    se_mean_height = z_sd_height / sqrt(s_number), #standard error of the sample mean
    z_mean_weight = mean(weight), #mean of sample height
    z_sd_weight = sd(weight), #sd of sample height
    se_mean_weight = z_sd_weight / sqrt(s_number), #standard error of the sample mean
    z_mean_age = mean(age), #mean of sample height
    z_sd_age = sd(age), #sd of sample height
    se_mean_age = z_sd_age / sqrt(s_number), #standard error of the sample mean
    z_mean_zombies_killed = mean(zombies_killed), #mean of sample height
    z_sd_zombies_killed = sd(zombies_killed), #sd of sample height
    se_mean_zombies_killed = z_sd_zombies_killed / sqrt(s_number), #standard error of the sample mean
    z_mean_years_of_eduction = mean(years_of_education), #mean of sample height
    z_sd_years_of_education = sd(years_of_education), #sd of sample height
    se_mean_years_of_education = z_sd_years_of_education / sqrt(s_number), #standard error of the sample mean
    
    
    #decade #adding decade column for inner_join in step 7 (whoops)
  )


#95% Confidence interval of Height
(ci_height <- sample_z$z_mean_height + c(-1, 1) * qnorm(1- 0.05) * sample_z$se_mean_height)

#95% Confidence interval of Wight
(ci_height <- sample_z$z_mean_weight + c(-1, 1) * qnorm(1- 0.05) * sample_z$se_mean_weight)

#95% Confidence interval of Age
(ci_height <- sample_z$z_mean_age + c(-1, 1) * qnorm(1- 0.05) * sample_z$se_mean_age)

#95% Confidence interval of zombies_killed
(ci_height <- sample_z$z_mean_zombies_killed + c(-1, 1) * qnorm(1- 0.05) * sample_z$se_mean_zombies_killed)

#95% Confidence interval of years_of_education
(ci_height <- sample_z$z_mean_years_of_education + c(-1, 1) * qnorm(1- 0.05) * sample_z$se_mean_years_of_education)
```

### Step 7

```{r}
#Running a sample distribution of mean height
#PReliminary
set.seed(123)
reps_z <- 199  # number of replications
n_per_sample <- 50 # number of values per replication
samp_dist_height <- vector("list", reps) #creating my dataframe as a vector list with my reps

#Bulk of code; thank you so much Jamie, I still keep getting weird squares but everything will hopefully look better on your end when you run without a packed environment :)
for (i in seq_len(reps_z)) {
  sampled_reps <- z %>% #designating reps to be pulled from z
   slice_sample(n = n_per_sample, replace = FALSE) #using splice_sample function again to pull my samples
  samp_dist_height[[i]] <- sampled_reps %>%
    summarise(
      z_mean_h = mean(height, na.rm = TRUE), #sample mean
      z_sd_h   = sd(height,   na.rm = TRUE) #sample sd
    )
}
samp_dist_height <- bind_rows(samp_dist_height)



#Running a sample distribution of mean weight
#PReliminary
set.seed(1234)
reps_z <- 199  # number of replications
n_per_sample <- 50 # number of values per replication
samp_dist_weight <- vector("list", reps) #creating my dataframe as a vector list with my reps

#Bulk of code
for (i in seq_len(reps_z)) {
  sampled_reps <- z %>% #designating reps to be pulled from z
   slice_sample(n = n_per_sample, replace = FALSE) #using splice_sample function again to pull my samples
  samp_dist_weight[[i]] <- sampled_reps %>%
    summarise(
      z_mean_w = mean(weight, na.rm = TRUE), #sample mean
      z_sd_w = sd(weight,   na.rm = TRUE) #sample sd
    )
}
samp_dist_weight <- bind_rows(samp_dist_weight)

#Running a sample distribution of mean age
#PReliminary
set.seed(1235)
reps_z <- 199  # number of replications
n_per_sample <- 50 # number of values per replication
samp_dist_age <- vector("list", reps) #creating my dataframe as a vector list with my reps

#Bulk of code
for (i in seq_len(reps_z)) {
  sampled_reps <- z %>% #designating reps to be pulled from z
   slice_sample(n = n_per_sample, replace = FALSE) #using splice_sample function again to pull my samples
  samp_dist_age[[i]] <- sampled_reps %>%
    summarise(
      z_mean_a = mean(age, na.rm = TRUE), #sample mean
      z_sd_a = sd(age,   na.rm = TRUE) #sample sd
    )
}
samp_dist_age <- bind_rows(samp_dist_age)


#Running a sample distribution of mean zombies killed
#PReliminary
set.seed(1236)
reps_z <- 199  # number of replications
n_per_sample <- 50 # number of values per replication
samp_dist_killedz <- vector("list", reps) #creating my dataframe as a vector list with my reps

#Bulk of code
for (i in seq_len(reps_z)) {
  sampled_reps <- z %>% #designating reps to be pulled from z
   slice_sample(n = n_per_sample, replace = FALSE) #using splice_sample function again to pull my samples
  samp_dist_killedz[[i]] <- sampled_reps %>%
    summarise(
      z_mean_zk = mean(zombies_killed, na.rm = TRUE), #sample mean
      z_sd_zk = sd(zombies_killed, na.rm = TRUE) #sample sd
    )
}
samp_dist_killedz <- bind_rows(samp_dist_killedz)


#Running a sample distribution of mean education
#PReliminary
set.seed(1237)
reps_z <- 199  # number of replications
n_per_sample <- 50 # number of values per replication
samp_dist_edu <- vector("list", reps) #creating my dataframe as a vector list with my reps

#Bulk of code
for (i in seq_len(reps_z)) {
  sampled_reps <- z %>% #designating reps to be pulled from z
   slice_sample(n = n_per_sample, replace = FALSE) #using splice_sample function again to pull my samples
  samp_dist_edu[[i]] <- sampled_reps %>%
    summarise(
      z_mean_edu = mean(years_of_education, na.rm = TRUE), #sample mean
      z_sd_edu = sd(years_of_education, na.rm = TRUE) #sample sd
    )
}
samp_dist_edu <- bind_rows(samp_dist_edu)

```

### Step 8

```{r}
#Plotting the sampling distributions for each variable mean with random f^$%ing colors

#Plotting height sampling distribution
ggplot(samp_dist_height, aes(x = z_mean_h)) +
  geom_histogram(binwidth = 10, color = "blue", fill = "green")
#Slight skew 

#Plotting the weight sampling distribution
ggplot(samp_dist_weight, aes(x = z_mean_w)) +
  geom_histogram(binwidth = 1, color = "pink", fill = "blue")
#Normal distribution 

#Plotting age sampling distribution
ggplot(samp_dist_age, aes(x = z_mean_a))
  geom_histogram(binwidth = 1, color = "blue", fill = "pink")
  
#PLotting zombie killed sampling distribution
ggplot(samp_dist_killedz, aes(x = z_mean_zk)) +
  geom_histogram(binwidth = 1, color = "violet", "pink")
  
#Plotting education sampling distribution
ggplot(samp_dist_edu, aes(x = z_mean_edu))
  geom_histogram(binwidth = 1, color = "pink", fill = "violet")
  
```

### Step 9

```{r}
#Creating 95% confidence intervals for each variable in the overal 'z' data

ci_95_height <- samp_dist_height %>%
  summarise(
    ci_lower = quantile(z_mean_h, probs = 0.025),
    ci_upper = quantile(z_mean_h, probs = 0.975),
    reps = n(),
    .groups = "drop"
  )
ci_95_height

ci_95_weight <- samp_dist_weight %>%
  summarise(
    ci_lower = quantile(z_mean_w, probs = 0.025), 
    ci_upper = quantile(z_mean_w, probs = 0.975),
    reps = n(),
    .groups = "drop"
  )
ci_95_weight

ci_95_age <- samp_dist_age %>%
  summarise(
    ci_lower = quantile(z_mean_a, probs = 0.025),
    ci_upper = quantile(z_mean_a, probs = 0.975),
    reps = n(),
    .groups = "drop"
  )
ci_95_age

ci_95_zkill <- samp_dist_killedz %>%
  summarise(
    ci_lower = quantile(z_mean_zk, probs = 0.025),
    ci_upper = quantile(z_mean_zk, probs = 0.975),
    reps = n(),
    .groups = "drop"
  )
ci_95_zkill

ci_95_edu <- samp_dist_edu %>%
  summarise(
    ci_lower = quantile(z_mean_edu, probs = 0.025),
    ci_upper = quantile(z_mean_edu, probs = 0.975),
    reps = n(),
    .groups = "drop"
  )
ci_95_edu
```

### Step 10

```{r}
#Generating 95% confidence intervals for bootstraps (made for walkin')
z_boot <- 10000

boot_height <- do(z_boot) * 
  quantile(boot_height, probs = c(0.025, 0.975), na.rm = TRUE)
  
(boot_weight <- do(z_boot) * 
 quantile(boot_weight, probs = c(0.025, 0.975), na.rm = TRUE))

boot_a <- do(z_boot) *
  quantile(boot_a, probs = c(0.025, 0.975), na.rm = TRUE)

boot_kill <- do(z_boot) *
  quantile(boot_kill, probs = c(0.025, 0.975, na.rm = TRUE))

boot_edu <- do(z_boot) *
  quantile(boot_edu, probs = c(0.025, 0.975), na.rm = TRUE)

#Due to high replicate amount, no large difference observed between confidence intervals
```
